Details matter: Similar-looking host galaxies can produce different merging compact binary populations

### 2015-05-29 paper/paper.tex
new edits yay

\subsection{Simulating galaxy evolution}
To thoroughly examine the significance of low metallicity star
formation, we have run a cosmological smoothed particle hydrodynamics
(SPH) $N$-body simulation of a Milky Way-like galaxy with GASOLINE
\citep{Stadel01,Wadsley04}.  This simulation allows us to analyze both
spatially and temporally resolved star formation, and determine not
only the metallicity history of compact object progenitors, but also
their dynamical evolution.  One simulation we report here, $h258$,
involves the formation of a Milky Way-like disk galaxy whose
progenitors undergo major mergers at $z = 2$ and $z = 0.8$.  This
simulation has been previously discussed by
\citet{Governato09,Bellovary10,Bellovary11,Brooks11}; however we have
re-run the simulation at a factor of two higher in spatial resolution
and additionally included new physics including a prescription for
cooling by molecular hydrogen (Christensen et al. in prep).

We selected our simulated region of interest from a 50 Mpc volume of
uniform resolution, and resampled the region of interest at very high
resolution using the volume renormalization technique \citep{Katz93}.
This technique allows us to follow the detailed physical processes
involved in galaxy evolution in our selected region while still
including large-scale torques from cosmic structure.  We use {\bf
WMAPX} cosmology \citep{WMAP} and model the ionizing UV background
with the prescription from \citet{Haardt96}.  Stars form
probabilistically from gas particles which meet density ($n_{min} =
10$ amu cm$^{-3}$) and temperature ($T_{max} = 10^4$ K) thresholds
{\bf (additional description needed for H2 SF prescription)}. If a gas
particle meets these criteria, it has a likelihood of forming a star
particle (representing a simple stellar population with a Kroupa IMF
\citep{Kroupa}) which is given by

\begin{equation}
p = \frac{m_{\rm gas}}{m_{\rm star}} (1 - e^{c^*\Delta t/t_{\rm form}})
\end{equation}

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{Kennicutt89}; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses;\footnote{In our high-resolution simulations \textbf{ROS-check!}, each gas particle starts with a mass
  $m_{gas,0}=26,676 M_\odot$, gaining mass from feedback and losing it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{2010ApJ...717..121C} for a discussion
of resolution issues in these SPH simulations.} $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  We model supernova feedback using the
blastwave formalism described in \citet{McKee77} and implemented in
our simulations as in \citet{Stinson06}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{Brooks07}, the Tully-Fisher relation \citep{Governato09}, and
the size-luminosity relation \citep{Brooks11}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{Governato10} and the Milky Way \citep{Guedes11}.

Also included in our simulations is a scheme for turbulent metal
diffusion \citep{Shen10}.  Metals are created in supernova explosions
and deposited directly to the gas within the blast radius.  Stellar
masses are converted to ages as described by \citet{Raiteri96}, and
stars more massive than 8 M$_\odot$ are able to undergo a Type II
supernova.  We follow metal enrichment from both Type II and Type 1a
supernova, with metal yields derived from \citet{Weaver93} and
\citet{Thielemann86} respectively.  From this point onwards metals
diffuse through the surrounding gas, according to 
\begin{equation}
\frac{dZ}{dt}|_{diff} = \nabla (D \nabla {Z}) \\
\end{equation}
%
where the diffusion paremeter $D$ is given by
%
\begin{equation}
D = C_{diff} |S_{ij}| h
\end{equation}
%
and $h$ is the SPH smoothing length, $S_{ij}$ is the trace-free shear
tensor, and $C_{diff}$ is a dimensionless constant which we set to
{\bf something.  Reasons for setting it are good, let's have some.
Sijing finds that C = 0.05 makes for a good comparison to clusters, we
tend to use something more conservative}.

We identify individual galaxies using the halo finder $AHF$
\citep{Gill04,Knollmann09}, which identifies halos based on an
overdensity criterion for a flat universe \citep{Gross97}.  For this
work, we are focusing on the primary (i.e. most massive) galaxy at any
given redshift, which serves as the progenitor to our $z = 0$ Milky
Way-like galaxy.

@@

{\bf Richard go ahead and edit this however is needed.}  The evolution
of a galaxy, in terms of its stellar mass and metallicity evolution,
depends strongly on its interaction history.  Galaxies which look
similar at the present day may have had drastically different
histories, which may result in differences in compact object merger
rates.  To investigate whether galaxy history affects the GRB event
rate, we have chosen two simulations which are extremely similar at $z
= 0$ but differ strongly in their merger histories.

The simulation h277 (Boring Galaxy) is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{Loebman12,Loebman14,Kassin14}, baryon fraction
\citep{Munshi13}, halo properties \citep{Zolotov09,Zolotov10}, and
  disk properties \citep{Brooks11}.  It has a virial mass of $M_{vir}
  = 6.79 \times 10^{11}  \msun$, stellar mass $M_* = 4.24 \times
  10^{10} \msun$, and circular velocity $v_{circ} = \editremark{xx}$ km s$^{-1}$.

  The simulation h258 (Exciting Galaxy), on the other hand, has a much
  more active merger history.  At $z = 1$ there is a 1:1 merger event;
  a gas disk rapidly reforms following the collision (see
  \citet{Governato09}), resulting in a massive disk galaxy at $z = 0$
  which looks remarkably similar to the Milky Way and to the other
  simulation, h277.  Prior to the $z = 1$ merger, each of the two
  progenitor galaxies actually experiences its own additional major
  merger events around $z = 2.5$ ({\bf check!}).  The combination of
  the series of major mergers, plus a number of minor interactions and
  flybys, gives a stark contrast to the relatively quiescent history
  of h277.  At $z = 0$, h258 has a virial mass of $M_{vir} = 7.74
  \times 10^{11} \msun$, stellar mass $M_* = 4.46 \times 10^{10} \msun$, and circular velocity $v_{circ} = \editremark{xx}$ km s$^{-1}$.

  Due to the differences in merger histories, the metallicity
  evolution of h258 and h277 also differ at early times.  Figure
  \ref{fig:TwoGalaxies} shows the history of the mean metallicty of
  recently formed stars vs time (thick solid lines), where we define
  recent as within the past 50 Myr.  The shaded regions correspond to
  one standard devation of the mean, while the thick dashed lines
  represent the top and bottom 90\%.  We see that {\bf different things happen at different times.}

>>>>

\subsection{Simulating galaxy evolution}
To thoroughly study the significance of low metallicity star
formation, we examine cosmological smoothed particle hydrodynamics
(SPH) $N$-body simulations of Milky Way-like galaxies with GASOLINE
\citep{2001PhDT........21S,2004NewA....9..137W}.  These simulations allow us to analyze both
spatially and temporally resolved star formation, and determine not
only the metallicity history of compact object progenitors, but also
their dynamical evolution.  
%maybe move the below section to later and/or consolidate
One simulation we report here, $h258$,
involves the formation of a Milky Way-like disk galaxy whose
progenitors undergo major mergers at $z = 2$ and $z = 0.8$.  This
simulation has been previously discussed by
\citet{2009MNRAS.398..312G,2010ApJ...721L.148B,2011ApJ...742...13B,2011ApJ...728...51B}; however we have
re-run the simulation at a factor of two higher in spatial resolution
and additionally included new physics including a prescription for
cooling by molecular hydrogen \citet{2012MNRAS.425.3058C}.

We selected our simulated region of interest from a 50 Mpc volume of
uniform resolution, and resampled the region at very high
resolution using the volume renormalization technique \citep{1993ApJ...412..455K}.
This technique allows us to follow the detailed physical processes
involved in galaxy evolution in our selected region while still
including large-scale torques from cosmic structure.  We use
WMAP3 cosmology \citep{WMAP} and model the ionizing UV background
with the prescription from \citet{1996ApJ...461...20H}.  Stars form
probabilistically from gas particles which meet density ($n_{min} =
10$ amu cm$^{-3}$) and temperature ($T_{max} = 10^4$ K) thresholds. If a gas
particle meets these criteria, it has a likelihood of forming a star
particle (representing a simple stellar population with a Kroupa IMF
\citep{2001MNRAS.322..231K}) which is given by

\begin{equation}
p = \frac{m_{\rm gas}}{m_{\rm star}} (1 - e^{c^*\rm X_{\rm H_2} \Delta t/t_{\rm form}})
\end{equation}

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{1989ApJ...344..685K};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses;\footnote{In our high-resolution simulations \textbf{ROS-check!}\editremark{this number is correct, is that what you want to check? - JMB}, each gas particle starts with a mass
  $m_{gas,0}=26,676 M_\odot$, gaining mass from feedback and losing it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{2010ApJ...717..121C} for a discussion
of resolution issues in these SPH simulations.} $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  We model supernova feedback using the
blastwave formalism described in \citet{1977ApJ...218..148M} and implemented in
our simulations as in \citet{2006MNRAS.373.1074S}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{2007ApJ...655L..17B}, the Tully-Fisher relation \citep{2009MNRAS.398..312G}, and
the size-luminosity relation \citep{2011ApJ...728...51B}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{2010Natur.463..203G} and the Milky Way \citep{2011ApJ...742...76G}.

Also included in our simulations is a scheme for turbulent metal
diffusion \citep{2010MNRAS.407.1581S}.  Metals are created in supernova explosions
and deposited directly to the gas within the blast radius.  Stellar
masses are converted to ages as described by \citet{1996A&A...315..105R}, and
stars more massive than 8 M$_\odot$ are able to undergo a Type II
supernova.  We follow metal enrichment from both Type II and Type 1a
supernova, with metal yields derived from \citet{1993PhR...227...65W} and
\citet{1986A&A...158...17T} respectively.  From this point onwards metals
diffuse through the surrounding gas, according to 
\begin{equation}
\frac{dZ}{dt}|_{diff} = \nabla (D \nabla {Z}) \\
\end{equation}
%
where the diffusion paremeter $D$ is given by
%
\begin{equation}
D = C_{diff} |S_{ij}| h
\end{equation}
%
and $h$ is the SPH smoothing length, $S_{ij}$ is the trace-free shear
tensor, and $C_{diff}$ is a dimensionless constant which we set to a conservative value of 0.03.

%Sijing finds that C = 0.05 makes for a good comparison to clusters, and that C = 0.05 - 0.1 is expected from turbulence theory.

We identify individual galaxies using the halo finder $AHF$
\citep{2004MNRAS.351..399G,2009ApJS..182..608K}, which identifies halos based on an
overdensity criterion for a flat universe \citep{1997PhDT........14G}.  For this
work, we are focusing on the primary (i.e. most massive) galaxy at any
given redshift, which serves as the progenitor to our $z = 0$ Milky
Way-like galaxy.

@@

{\bf Richard go ahead and edit this however is needed.}  The evolution
of a galaxy, in terms of its stellar mass and metallicity evolution,
depends strongly on its interaction history.  Galaxies which appear
similar at the present day may have had drastically different
histories, which may result in differences in compact object merger
rates.  To investigate whether galaxy history affects the GRB event
rate, we have chosen two simulations which are extremely similar at $z
= 0$ but differ strongly in their merger histories.

The simulation h277 (Boring Galaxy) is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{2012ApJ...758L..23L,2014ApJ...794..151L,2014ApJ...790...89K}, baryon fraction
\citep{2013ApJ...766...56M}, halo properties \citep{2009ApJ...702.1058Z,2010ApJ...721..738Z}, and
disk properties \citep{2011ApJ...728...51B}.  It has a virial mass of $M_{vir} =
6.79 \times 10^{11} \msun$, stellar mass $M_* = 4.24 \times 10^{10}
\msun$, and maximum circular velocity $v_{circ} = 235$ km s$^{-1}$.

  The simulation h258 (Exciting Galaxy), on the other hand, has a much
  more active merger history.  At $z = 1$ there is a 1:1 merger event;
  a gas disk rapidly reforms following the collision (see
  \citet{2009MNRAS.398..312G}), resulting in a massive disk galaxy at $z = 0$
  which looks remarkably similar to the Milky Way and to the other
  simulation, h277.  Prior to the $z = 1$ merger, each of the two
  progenitor galaxies actually experiences its own additional major
  merger events around $z = 2$.  The combination of
  the series of major mergers, plus a number of minor interactions and
  flybys, gives a stark contrast to the relatively quiescent history
  of h277.  At $z = 0$, h258 has a virial mass of $M_{vir} = 7.74
  \times 10^{11} \msun$, stellar mass $M_* = 4.46 \times 10^{10}
  \msun$, and maximum circular velocity $v_{circ} = 242$ km s$^{-1}$.  \editremark{Might we want to show the images of these galaxies?}

  Due to the differences in merger histories, the metallicity
  evolution of h258 and h277 also differ at early times.  Figure
  \ref{fig:TwoGalaxies} shows the history of the mean metallicty of
  recently formed stars vs time (thick solid lines), where we define
  recent as within the past 50 Myr.  The shaded regions correspond to
  one standard devation of the mean, while the thick dashed lines
  represent the top and bottom 90\%.  We see that the early evolution
  of the metal properties of these galaxies does differ - between 0.5
  and 6 Gyr, h277 hosts a {\bf modestly} more metal-rich population than
  h258. \editremark{The difference isn't all that big, I want to address it without sounding foolish.}
